---
title: "Lab 2 - Introduction to R"
author: "ECOL 620 - Applications in Landscape Ecology"
date: "Spring 2023"
output: 
  html_document:
    toc: true
    toc_float: true
    # code_folding: hide
editor_options: 
  chunk_output_type: console
knit: (function(inputFile, encoding){ 
    out_dir <- '../';
    rmarkdown::render(inputFile, encoding = encoding, output_file=file.path(dirname(inputFile), out_dir, '/data/lab2_practice.html')) 
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Through this lab we will explore core concepts of scale using R. Here, we focus on spatial scale, i.e., grain and extent, and examine how these properties manifests in ecological conclusions and inference. Download laboratory #2 materials and store files locally on your computer.


```{r}
#load packages
library(terra)
library(tidyverse)
```


A few simple example of rasters
```{r}
set.seed(16)
#xmx maximum x coordinate (right border). Modify the number to 10 and see what happens 
toy = rast(ncol=6, nrow=6, xmin=0, xmax=6, ymin=0, ymax=6)
toy[] = rpois(ncell(toy), lambda=3)
```

```{r}
#plot
plot(toy, axes=F)
text(toy, digit=2)
```

```{r}
res(toy) #resolution of the raster, i.e. grain 1 by 1
```


```{r}
#check cell labeling/order
ncell(toy) #number of cells
toy2 = toy
toy2[] = 1:ncell(toy)
```


```{r}
#plot
plot(toy2)
text(toy2, digits=2)
```


```{r}
#increase the grain
toy_mean = aggregate(toy, fact=2, fun="mean") #mean value
toy_maj = aggregate(toy, fact=2, fun="modal") #majority rule
```

```{r}
#plot mean rule
plot(toy_mean)
text(toy_mean,digits=1)
```


```{r}
#plot majority rule
plot(toy_maj)
text(toy_maj)
```


```{r}
#contrast means/standard deviation
global(toy, "mean")
global(toy, fun="sd")
```

```{r}
#Note the reduction in sd, relative to the sd of toy
global(toy_mean, "mean")
global(toy_mean, "sd")
```


```{r}
global(toy_maj, "mean")
global(toy_maj, "sd")
```


```{r}
#decrease the grain
toy_dis2 = disagg(toy, fact=2) #simple method, increase by a factor of 2 without any interpolation
```


```{r}
toy_dis2_bilinear = disagg(toy, fact=2, method='bilinear') # increase by a factor of 2 using bilinear interpolation
```


```{r}
#plot
plot(toy_dis2, axes=F)
plot(as.polygons(toy_dis2, dissolve=F), add=TRUE, border='gray50', lwd=1)
text(toy_dis2, cex=0.9)
```

```{r}
#plot
plot(toy_dis2_bilinear, axes=F)
plot(as.polygons(toy_dis2_bilinear, dissolve=F), add=TRUE, border='gray50', lwd=1)
text(toy_dis2_bilinear, digits=1, cex=0.6)
```

```{r}
#decrease the extent
e = ext(2, 4, 2, 4)#first create new, smaller extent
toy_crop = crop(toy, e)
```


```{r}
#plot
plot(toy, range=c(0,7), type = 'continuous')
rect(2, 2, 4, 4, border = "black", lwd = 2)
plot(toy_crop, range=c(0,7), type = 'continuous')
```

```{r}
#increase the extent
e = ext(0, 7, 0, 7)#first create new, bigger extent
toy_big = extend(toy,e)
```

```{r}
#plot
plot(toy)
plot(toy_big)
```
```{r}
remove(toy,toy2, e, toy_big, toy_crop, toy_dis2, toy_dis2_bilinear, toy_maj, toy_mean)
```


####################################
#2.3.4.1 multi-scale analysis
####################################

#------------------#
#nlcd
#------------------#
```{r}

nlcd<-rast("../data/nlcd2011SE/nlcd2011SE.tif")
```

```{r}
#inspect
crs(nlcd) #alternative function from terra package
```

```{r}
#set projection
nlcd_proj <- crs(nlcd)
```


```{r}
#inspect raster properties
res(nlcd)
ncell(nlcd)
ext(nlcd)
```

```{r}
#check raster values
levels(nlcd)
nlcd <- as.factor(nlcd) #convert to factors, this may take a little while... (~1-2 mintues)
levels(nlcd)
plot(nlcd)
```


#-------------------------------#
#site locations: shp file
#-------------------------------#

```{r}
library(sf)
```

```{r}
#site and reptile data
sites = st_read("../data/reptiledata/reptiledata.shp")
```

```{r}
#inspect
class(sites)
```

```{r}
st_crs(sites)
```

```{r}
st_crs(sites) = nlcd_proj #set projection
summary(sites)
```


```{r}
head(sites, 2)
```

```{r}
#plot with custom color scheme
my_col = c("black","blue","darkorange","red","darkred","grey30","grey50", "lightgreen",
            "green", "darkgreen", "yellow", "goldenrod", "purple", "orchid","lightblue", "lightcyan")

```



```{r, warning=F}
#plot
plot(nlcd, col=my_col)
plot(sites, add=T, col="white")

```

```{r}
#subset points to remove corn land use
sites = sites %>% filter(management!="Corn")
nrow(sites)
```

```{r}
#crop raster to 10 km from sampling points: determine min/max coordinates for new extent
x.min = min(sites$coords_x1) - 10000
x.max = max(sites$coords_x1) + 10000
y.min = min(sites$coords_x2) - 10000
y.max = max(sites$coords_x2) + 10000
```


```{r}
extent.new <- ext(x.min, x.max, y.min, y.max)
nlcd = crop(nlcd, extent.new) #this may take ~20 seconds
```


```{r}
#reclassify with classify function is faster
levels(nlcd)[[1]]
reclass = c(rep(0,7), rep(1,3), rep(0,6))
nlcd.levels = levels(nlcd)[[1]]

#create reclassify matrix: first col: original; second: change to
reclass.mat = cbind(levels(nlcd)[[1]], reclass)
reclass.mat

#reclassify
# forest = classify(nlcd, reclass.mat[,2:3])
forest = classify(nlcd, reclass.mat[,1:2])
```

```{r, warning=F}
#plot
plot(forest)
plot(sites, pch=21, col="black", add=T)
```

```{r}
#define the buffer width (i.e., radius)
buf1km = 1000
buf5km = 5000

#buffer site 1
buffer.site1.1km = st_buffer(sites[1,], dist=buf1km, nQuadSegs = 30) #1km
buffer.site1.5km = st_buffer(sites[1,], dist=buf5km, nQuadSegs = 30) #5km

#buffer all sites
buffer.1km = st_buffer(sites, dist=buf1km, nQuadSegs = 30) #1km
buffer.5km = st_buffer(sites, dist=buf5km, nQuadSegs = 30) #5km
```



```{r, warning=F}
#don't worry, ggplot can used!
gg_forest=crop(forest, ext(buffer.site1.5km))
ggplot() +
  # geom_tile(data = as.data.frame(gg_forest, xy=T), aes(x=x,y=y, fill=as.factor(label))) +
  geom_tile(data = as.data.frame(gg_forest, xy=T), aes(x=x,y=y, fill=as.factor(nlcd2011SE))) +
  scale_fill_manual(values=c("white", "darkgreen"))+
   geom_sf(data = buffer.site1.5km,  fill = NA, color ="black", lwd=.1)+
  coord_sf(xlim=c(ext(buffer.site1.5km)[1],ext(buffer.site1.5km)[2]),
                  ylim=c(ext(buffer.site1.5km)[3],ext(buffer.site1.5km)[4]))+
  labs(fill="Cover Type")
```

```{r, warning=F}
#zoom in on plot for 5 km buffer at site 1
#can provide object to zoom on or click twice on layer

#plot
zoom(nlcd, buffer.site1.5km, col=my_col)
plot(buffer.site1.1km, border="red",col=NA, lwd = 3, add=T)
plot(buffer.site1.5km, border="red", col=NA, lwd = 3, add=T)
plot(sites[1,], col="grey20", bg="black", pch=22, cex=1, add=T)
```


```{r}
#calculate forest area within buffer
buffer.forest1.1km <- crop(forest, buffer.site1.1km)
buffer.forest1.1km <- mask(buffer.forest1.1km, buffer.site1.1km)

```

```{r}
#plot forest within buffer
plot(buffer.forest1.1km)
```


```{r}
#calculate percent forest cover
grainarea <- res(forest)[[1]]^2/10000#in ha
bufferarea <- (3.14159*buf1km^2)/10000#pi*r^2
forestcover1km <- global(buffer.forest1.1km, 'sum', na.rm=T)*grainarea
percentforest1km <- forestcover1km/bufferarea*100
percentforest1km
```



#-----------------------------------------#
#Function that puts all the steps together
#requires:
#  points: one set of x,y coordinates
#  size: the buffer size (radius), in m
#  landcover: a binary land-cover map
#  grain: the resolution of the map
#-----------------------------------------#

```{r}
BufferCover <- function(coords, size, landcover, grain){
  
  bufferarea.i <- pi*size^2/10000                             #size must be in m
  coords.i <- st_as_sf(as.data.frame(cbind(
    lon=coords$coords_x1[i],lat=coords$coords_x2[i])), 
    coords = c("lon", "lat"))                                #create spatial points from coordinates
  buffer.i = st_buffer(coords.i, dist=size)                   #buffer from rgeos
  crop.i = crop(landcover, buffer.i)                         #crop with raster function
  crop.NA = setValues(crop.i, NA)                            #empty raster for the rasterization
  buffer.r = rasterize(buffer.i, crop.NA)                    #rasterize buffer
  land.buffer = mask(x=crop.i, mask=buffer.r)                #mask by putting NA outside the boundary
  coveramount = global(land.buffer, 'sum', na.rm=T)*grain            #calculate area
  percentcover<-100*(coveramount/bufferarea.i)                #convert to %
  
  return(percentcover)
}
```

```{r}
#create empty vector for storing output
f1km <- rep(NA, length = nrow(sites))
f2km <- rep(NA, length = nrow(sites))
```

```{r}
#with for loop (all five buffers: 910s; <=3km: 228s)
xxx for(i in 1:nrow(sites)) {
  f1km[i] <- BufferCover(coords=sites,size=1000,landcover=forest,grain=grainarea)
  f2km[i] <- BufferCover(coords=sites,size=2000,landcover=forest,grain=grainarea)
  print(i)
}
```


```{r}
#make a data frame
forest.scale <- data.frame(site=sites$site,
                         x=sites$coords_x1, y=sites$coords_x2,
                         f1km=unlist(f1km), f2km=unlist(f2km))

```

```{r}
#plot
plot(f1km, f2km)
```

```{r}
#correlation matrix
cor(forest.scale[,4:5])
```



####################################
#2.3.4.2 Scale of effect
####################################

#----------------------------------------#
#2.3.4.2 Buffer analysis
#----------------------------------------#
```{r}
#herp data
flsk <- read.csv("../data/reptiledata/reptiles_flsk.csv", header=T)
flsk <- merge(flsk, forest.scale, by="site", all=F)

```

```{r}
#glms at 2 scales; see text for more scales considered
pres.1km <- glm(pres ~ f1km, family = "binomial", data = flsk)
pres.2km <- glm(pres ~ f2km, family = "binomial", data = flsk)
```

```{r}
#summary information
summary(pres.1km)
summary(pres.2km)
```


```{r}
#likelihoods
logLik(pres.1km)
logLik(pres.2km)
```

```{r}
#accessing coefficients
pres.1km.ci <- confint(pres.1km)
pres.2km.ci <- confint(pres.2km)
pres.1km.ci
pres.2km.ci
```




